### 2023-03-27 paper/paper.tex
update paper

\pdfoutput=1

\documentclass[11pt]{article}

\usepackage[]{acl}
\usepackage{tabularx}

\usepackage{times}
\usepackage{latexsym}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{microtype}

\title{Examining readability of text generated through GPT-3 }

\author{Aron Winkler \\
  University of Tübungen \\
  \texttt{winkler.aron5@gmail.com}}

\begin{document}
\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
\citep{corenlp} hello
\section{Dataset and metrics}

\subsection{Dataset}

\subsection{Metrics}

\section{Results}

\section{Discussion and future research}

\section*{Acknowledgements}

\bibliography{anthology,custom}

\appendix
\section*{Resources}

\label{sec:appendix}

All code for this paper is available at \url{https://github.com/waron97/nlp_for_readability_term_paper}.

\end{document}