### 2023-03-28 paper/paper.tex
update paper

\pdfoutput=1

\documentclass[11pt]{article}

\usepackage{graphicx}

\usepackage[]{acl}
\usepackage{tabularx}

\usepackage{times}
\usepackage{latexsym}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{microtype}

\title{Examining readability of text generated through GPT-3 }

\author{Aron Winkler \\
  University of Tübungen \\
  \texttt{winkler.aron5@gmail.com}}

\begin{document}
\maketitle
\begin{abstract}
    This paper examines readability-related metrics of AI-generated text, specifically texts produced by OpenAI's GPT-3. Generation was performed for the English language across 10 topics and the 6 proficiency levels defined by the CEFR framework, and outputs were compared with Simple Wikipedia and (standard) Wikipedia articles on the same topics. 7 commonly used complexity metrics have been employed, targeting chiefly lexis and syntax. In the limited scope of this experiment, GPT-3 is observed to at least partially aware of what constitutes reading difficulty.
\end{abstract}

\section{Introduction}

AI text generation has experiences a massive rise in popularity with the media reach achieved by the release of OpenAI's GPT-3 \citep{brown2020language} and its chat GUI interface ChatGPT\footnote{\url{https://openai.com/blog/chatgpt}}. While early public opinion focused on how - in the school setting - it would be students who would benefit from the new system by having their schoolwork done for them, little was said about how AI text generation could become an asset for didactitians in generating teaching material.

AI text generation might become an important tool for teachers, allowing them to generate learning material about topics and events that engage their students that might not be available on the market. An interesting setting for such a phenomenon is language learning, where it is necessary to produce learning material for students at different proficiency levels: whether to create ex-novo or simplify existing text material, AI text-generation tools can easily be imagined as having an enabling effect.

It is however not obvious that AI generation tools reach the desired quality or adherence to the didactic goals that teachers set out. Owing to the increased popularity of these tools, more research is desirable on their applicability in a number of different settings. For example, GPT-3 has been observed to "make up" information about some real-world events or phenomena, a fact that makes it hard to rely on for teaching. In the specific context of this experiment, employing GPT-3 for language-learning material is only valuable insofar as the generated content is adequate in difficulty to the audience.

This work aims to make a contribution to these questions by verifying the readability level of text content generated by GPT-3 with regard to 10 topics and the 6 proficiency levels defined by the CEFR framework. A number of complexity metrics are computed on the generated texts and compared with Simple Wikipedia\footnote{\url{https://simple.wikipedia.org/wiki/Main_Page}} as well as Wikipedia\footnote{\url{https://www.wikipedia.org/}} articles on the same topics in order to gauge the true readability levels of the model outputs.
\section{Dataset and metrics}
\subsection{Dataset}

OpenAI's REST api was employed to produce texts with the "text-davinci-003" model (a GPT-3 variant) across 10 topics\footnote{'Color Blindness, The Great Depression, Butterflies, Dogs, Semantics, The Internet, The Moon, Dinosaurs, Economics, Quantum Mechanics'} and 6 CEFR proficiency levels\footnote{A1, A2, B1, B2, C1, C2}. Topics were selected without any particular guiding principle other than a general goal of variety between events, entities, and phenomena. The prompts fed to the model followed the template:

\begin{quote}
  \emph{Write a text for learners of English at the [level] level about [topic].}
\end{quote}

In addition to the 60 texts obtained in this way, the dataset was enriched by Simple Wikipedia and Wikipedia articles corresponding to each of the 10 topics, resulting in an additional 20 dataset entries.

\subsection{Preprocessing}

Preprocessing of the dataset comprised enrichment with POS and dependency annotation. Part-of-speech tagging was achieved with nltk \citep{LoperBird02} and dependency relations were added through nltk in conjunction with CoreNLP models \citep{corenlp}.

\subsection{Metrics}

\input{tables/metrics.tex}

Table~\ref{tab:metrics} lists the readability metrics used in this experiment. This specific selection was inspired by \citep{venturi-etal-2015-nlp}, but many more examples of their usage can be found.

This feature set almost exclusively targets lexis and syntax, with little regard for morphology outside of the noun-to-verb ratio feature. While this is a limitation of this set, the language of this experiment is solely English, a language not necessarily known for a complex morphological system.

The feautre "Words in Top 3000 English words" uses unigram frequencies from the Google Web Trillion Word corpus\footnote{\url{https://ai.googleblog.com/2006/08/all-our-n-gram-are-belong-to-you.html}} to isolate the 3000 most frequent words of the English language. Many research approaches to readability employ similar ideas to gain insight into the lexical complexity of texts, such as the BIV (basic Italian vocabulary) employed by \citep{dellorletta-etal-2011-read}.

All feature values outside of type-token-ratio (which ignores sentence boundaries) are computed at the sentence level. When a dataset entry comprises multiple sentences, the output for that entry is the average of the values for the sentences that compose it.

\section{Results}

\input{tables/means.tex}

Table~\ref{tab:means} details the average output for each difficulty level and metric. For each metric, the 3 values most indicative of low readability are highlighted.

With regards to the selected metrics, GPT-3 does a good job generating texts for the target readability level in the sense that prompts requesting C1 and C2 level material on average yield the least readable texts, whereas lower proficiency levels tend to have scores indicating higher readability.

While no rigorous checks are carried out in this paper in reference to the correlation between these metrics and article readability levels, it's still possible to examine some of the results to gain some insight into the black box that is GPT-3.

\input{figs/wit_means.tex}

Among the selected measures, WIT (ratio of words in the 3000 most frequent English words) appears to be the most well-behaved, insofar as the prompted CEFR levels mostly maintain their natural progression. Figure~\ref{fig:wit-mean} shows the isolated scores for the WIT metric, which show a correlation with difficulty. Wikipedia and Simple Wikipedia articles have lower values, however this is possibly connected to those texts being several times longer.

\input{figs/lldl_means.tex}

LLDL (length of longest dependency link) is another metric that performs well in terms of level differentiation. It also reflects the common belief that the biggest difficulty spike in language learning - as shown by Figure~\ref{fig:lldl-mean}. As with most metrics, however, the natural progression of the CEFR levels is lost, particularly in the middle portions.

Interesting discussions can be had around topics as well, since, depending on GPT-3's ability to generalize across domains, certain areas might be problem areas for it. For example, "semantics" and "quantum mechanics" would be hard to write about for A1 level students as they are likely to be children who have not interacted with those fields yet.

Generally, a difference in values is visible across the board, more so at the lower difficuly levels. At CEFR levels B2 and above, many of the metrics appear similar regardless of topic. Among the selected measures, NVR (noun to verb ratio) shows the highest cross-topic variation, maintaining strong differentitation even at the C2-level texts. Figure \ref{fig:wit-nvr-topics} shows metric values at the A1 and C2 levels, which indicate that topics commonly regarded as intrinsically have higher information density in GPT-3 texts. This is however a weak case, as few metrics differ so sharply as NVR, and bars in these figures correspond to a single text entry.

\input{figs/nvr_topics.tex}

\section{Discussion and future research}

This work attempted to gauge true readability levels of GPT-3 output texts with prompts across 10 topics and CEFR proficiency levels, optimising thus for the second language learning application environment. Outputs computed from the metrics set suggest that GPT-3 is at least somewhat aware of syntax and lexis when performin generation for a target readability level.

Future research into this topic could explore a wider variety of textual metrics, as this work only selected a handful that lent themselves particularly well to interpretation and did not require steep hardware requirements to obtain.

Since recent research (e.g. \citealp{Deutsch_2020}) has also focused on readability assessment through neural approaches without such metrics, this could also be explored with AI text data. An attempt was made in relationship to this experiment to train a readability analyzer with GLoVe word embeddings \citep{pennington2014glove} and Wikipedia data, but this unfortunaly was inconclusive, and thus results from this analyzer haven't been reported.

Mainly, however, future extensions of this experiment or related experiments would benefit from higher volume. Simply put, more data is necessary across multiple domains to make educated conclusions about the quality of GPT-3 outputs in a teaching setting. While data is easy to generate, quality annotations on it are yet unavailable. Indeed, it is a good question whether making any is a fruitful enterprise, as the actual models behind GPT-3 and ChatGPT are constantly updated - thus making it hard to pinpoint what exactly is being studied.

\section*{Acknowledgements}

\bibliography{anthology,custom}

\appendix
\section*{Resources}

\label{sec:appendix}

All code for this paper is available at \url{https://github.com/waron97/nlp_for_readability_term_paper}.

\end{document}